
# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data
Read in the CSV file and classify columns.
```{r Readin, echo=TRUE}
activity <- read.csv("activity.csv", colClasses = c("numeric","Date","numeric"))
```
## What is mean total number of steps taken per day?
1. Create a Histogram Utilizing ggplot2.  Plot total number of steps taken per day.

```{r HistPlot, echo=TRUE}
library(ggplot2)
p <- ggplot(activity, aes(x=date, y=steps)) + geom_histogram(stat='identity', position='dodge')
datebreaks <- seq(as.Date("2012-10-01"), as.Date("2012-11-30"), by="1 week")
p 
```

2. Calculate and report the mean and median total number of steps taken per day

```{r MeanCalc, echo=TRUE}
library(doBy)
summaryBy(steps ~ date, data = activity, 
          FUN = list(mean, max, min, median, sd))
```

## What is the average daily activity pattern?
1. Create as Time Series plot of of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).

```{r Plot1, echo=TRUE}
library(plyr)
activity2 <- na.omit(activity)
activity3 <- ddply(activity2, c("interval"), na.rm = FALSE, summarise, avg = mean(steps))
plot(activity3$interval, activity3$avg, type="l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

A.  Maximun number of steps is:
```{r Interval1, echo=TRUE}
max(activity3$avg)
```
B. The 5-minute interval with the maximun number of steps is:
```{r Interval2, echo=TRUE}
activity3 <- activity3[order(-activity3$avg),]
activity3[1, ]
```

## Imputing missing values

1.  Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r MissingVals, echo=TRUE}
length(activity[is.na(activity)])
```

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

This code chunk will replace NA's with the mean.
```{r MissingVals2, echo=TRUE}
library(plyr)
impute.med <- function(steps) replace(steps, is.na(steps), mean(steps, na.rm = TRUE))
dat2 <- sapply(activity, function(steps){
    if(is.numeric(steps)){
            impute.med(steps)
        } else {
            steps
        }
    }
)
```

This code chunk will create a data frame with replaced NA's and binds the date.
```{r MissingVals3, echo=TRUE}
library(data.table)
activity2 <- data.frame(dat2)

activity2a <- activity2[c(1, 3)]
activitya <- activity[c(2)]

activity3 <- cbind(activity2a, activitya)

p <- ggplot(activity3, aes(x=date, y=steps)) + geom_histogram(stat='identity', position='dodge')
p 

library(doBy)
summaryBy(steps ~ date, data = activity3, 
          FUN = list(mean, median))
```          
## Are there differences in activity patterns between weekdays and weekends?

```{r Weekdays, echo=TRUE}
activity3$day <- weekdays(activity3$date)

activity3 = within(activity3,
{
        Week = as.factor( 
                ifelse(
                        (activity3$day) 
                        %in% c("Saturday","Sunday") 
                , "weekend", "weekday"
        )
)
}
)
```

Subset the data into two data frames.  One for weekends and one for weekdays.

```{r Weekdays2, echo=TRUE}
weekend <- subset(activity3, day=="Saturday" | day=="Sunday")
weekday <- subset(activity3, day=="Monday" | day=="Tuesday" | day=="Wednesday" | day=="Thursday" | day == "Friday")
```

Plot the weekend and weekday data frames for comparison.

```{r Weekdays3, echo=TRUE}                  
library(plyr)

weekday2 <- ddply(weekday, c("interval"), na.rm = FALSE, summarise, avg = mean(steps))
plot(weekday2$interval, weekday2$avg, type="l")

weekend2 <- ddply(weekend, c("interval"), na.rm = FALSE, summarise, avg = mean(steps))
plot(weekend2$interval, weekend2$avg, type="l")

```
